\textbf{SPECFEM3D Cartesian}\\ \textbf{User Manual}

### 2014-09-16 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex
added doc/SEM_class_from_Yvon_Maday_to_Dimitri_Komatitsch_and_Jean_Pierre_Vilotte_04apr1995.pdf

\noindent The SPECFEM3D Cartesian package was first developed by Dimitri
Komatitsch and Jean-Pierre Vilotte at Institut de Physique du Globe
(IPGP) in Paris, France from 1995 to 1997 and then by Dimitri Komatitsch
and Jeroen Tromp at Harvard University and Caltech, USA, starting in 1998.
The story started on March 28, 1995, when Prof. Yvon Maday from CNRS and University of Paris, France, gave a lecture to
Dimitri Komatitsch and Jean-Pierre Vilotte at IPG about the nice properties of the Legendre spectral-element method with diagonal mass matrix that he had used for
other equations. We are deeply indebted and thankful to him for that.
That followed a visit by Dimitri Komatitsch to OGS (Istituto Nazionale di Oceanografia e di Geofisica Sperimentale) in Trieste, Italy, in February 1995
to meet with G\'eza Seriani and Enrico Priolo, who introduced him to their 2D Chebyshev version of the spectral-element method with a non-diagonal mass matrix.
We are deeply indebted and thankful to them for that.

@@

To visualize these files, we use an auxilliary program \texttt{combine\_vol\_data.f90}
to combine the data from all slices into one mesh file. To compile
it in the root directory, type:
\begin{lyxcode}
{\footnotesize make~xcombine\_vol\_data~}{\footnotesize \par}
\end{lyxcode}
which will create the executable \texttt{xcombine\_vol\_data} in the
directory \texttt{bin/}. To output the usage of this executable, type
`./bin/xcombine\_vol\_data` without arguments. As an example, to run
the executable you would use
\begin{lyxcode}
{\footnotesize cd~bin/~}{\footnotesize \par}

~{\footnotesize ./xcombine\_vol\_data~0~3~velocity\_Z\_it000400~../OUTPUT\_FILES/DATABASES\_MPI~../OUTPUT\_FILES~0~}{\footnotesize \par}
\end{lyxcode}
to create a low-resolution mesh file. The output mesh file will have
the name \texttt{velocity\_Z\_it000400.mesh}. We next convert the
\texttt{.mesh} file into the VTU (Unstructured grid file) format which
can be viewed in ParaView. For this task, you can use and modify the
script \texttt{mesh2vtu.pl} located in directory \texttt{utils/Visualization/Paraview/},
for example:
\begin{lyxcode}
{\footnotesize mesh2vtu.pl~-i~velocity\_Z\_it000400.mesh~-o~velocity\_Z\_it000400.vtu}{\footnotesize \par}
\end{lyxcode}
Notice that this Perl script uses a program \texttt{mesh2vtu} in the
\texttt{utils/Visualization/Paraview/mesh2vtu} directory, which further
uses the VTK \urlwithparentheses{http://www.vtk.org/} run-time library
for its execution. Therefore, make sure you have them properly set
in the script according to your system.

>>>>

\noindent The SPECFEM3D Cartesian package was first developed by Dimitri
Komatitsch and Jean-Pierre Vilotte at Institut de Physique du Globe
(IPGP) in Paris, France from 1995 to 1997 and then by Dimitri Komatitsch
and Jeroen Tromp at Harvard University and Caltech, USA, starting in 1998.
The story started on April 4, 1995, when Prof. Yvon Maday from CNRS and University of Paris, France, gave a lecture to
Dimitri Komatitsch and Jean-Pierre Vilotte at IPG about the nice properties of the Legendre spectral-element method with diagonal mass matrix that he had used for
other equations. We are deeply indebted and thankful to him for that.
That followed a visit by Dimitri Komatitsch to OGS (Istituto Nazionale di Oceanografia e di Geofisica Sperimentale) in Trieste, Italy, in February 1995
to meet with G\'eza Seriani and Enrico Priolo, who introduced him to their 2D Chebyshev version of the spectral-element method with a non-diagonal mass matrix.
We are deeply indebted and thankful to them for that.

@@

To visualize these files, we use an auxilliary program \texttt{combine\_vol\_data.f90}
to combine the data from all slices into one mesh file. To compile
it in the root directory, type:
\begin{lyxcode}
{\footnotesize make~xcombine\_vol\_data~}{\footnotesize \par}
\end{lyxcode}
which will create the executable \texttt{xcombine\_vol\_data} in the
directory \texttt{bin/}. To output the usage of this executable, type
\begin{lyxcode}
{\footnotesize ./bin/xcombine\_vol\_data~}{\footnotesize \par}
\end{lyxcode}
without arguments.

\texttt{xcombine\_vol\_data} will combine the data on the different processors (located in \texttt{OUTPUT\_FILES/DATABASES\_MPI}), for a given quantity and a given iteration, to one file. For example, if you want to combine \texttt{velocity\_Z}, for the iteration 400 on 4 processors, i.e. if you want to combine these files :

\texttt{proc000000\_velocity\_Z\_it000400.bin }

\texttt{proc000001\_velocity\_Z\_it000400.bin }

\texttt{proc000002\_velocity\_Z\_it000400.bin }

\texttt{proc000003\_velocity\_Z\_it000400.bin}

You have to go in the directory of the concerned example (where are the directories \texttt{DATA}, \texttt{OUTPUT\_FILES}, etc.).
Then, you can launch \texttt{xcombine\_vol\_data}, specifying the path where it is located. As an example, if the \texttt{DATA} and \texttt{OUTPUT\_FILES} directories of the concerned example are in the root specfem3d directory, \texttt{xcombine\_vol\_data} is in \texttt{./bin}, and you have to type:

\begin{lyxcode}
~{\footnotesize ./bin/xcombine\_vol\_data~0~3~velocity\_Z\_it000400~./OUTPUT\_FILES/DATABASES\_MPI~./OUTPUT\_FILES~0~}{\footnotesize \par}
\end{lyxcode}

Here, $0$ is the number of the first processor, $3$ the number of the last one, \texttt{velocity\_Z\_it000400} the name of the files we want to combine without the prefix \texttt{proc000000*\_}, \texttt{./OUTPUT\_FILES/DATABASES\_MPI} the directory where the files are located, \texttt{./OUTPUT\_FILES} the directory where the combined file will be stored, and $0$ is the parameter to create a low-resolution mesh file ($1$ for a high-resolution).

At the beginning of the source file \texttt{combine\_vol\_data.f90} (located in \texttt{src/auxiliaries/}), there is a logical constant parameter \texttt{USE\_VTK\_OUTPUT} whose default value is \texttt{.true.}, that which will provide the output combined file directly in the VTK format and you can use Paraview to visualize it.

If the value of this parameter is \texttt{.false.}, the output mesh file will have the name \texttt{velocity\_Z\_it000400.mesh}. Then we next have to convert the
\texttt{.mesh} file into the VTU (Unstructured grid file) format which can be viewed in ParaView. For this task, you can use and modify the
script \texttt{mesh2vtu.pl} located in directory \texttt{utils/Visualization/Paraview/}, for example:
\begin{lyxcode}
{\footnotesize mesh2vtu.pl~-i~velocity\_Z\_it000400.mesh~-o~velocity\_Z\_it000400.vtu}{\footnotesize \par}
\end{lyxcode}
Notice that this Perl script uses a program \texttt{mesh2vtu} in the \texttt{utils/Visualization/Paraview/mesh2vtu} directory, which further
uses the VTK \urlwithparentheses{http://www.vtk.org/} run-time library for its execution. Therefore, make sure you have them properly set
in the script according to your system.

Then, to do a movie with several iterations, you have to repeat this process for each iteration you want to put in your movie.
